### 2025-07-01 mixed-models.qmd
20250701 - standardized regression coefficients

```{r}
library("lme4")
library("lmerTest")
library("MuMIn")
library("emmeans")
library("sjstats")
library("mgcv")
library("AICcmodavg")
library("bbmle")
library("rstan")
library("brms")
library("cmdstanr") # todo: install.packages("cmdstanr", repos = c("https://stan-dev.r-universe.dev", getOption("repos"))); cmdstanr::check_cmdstan_toolchain(); cmdstanr::install_cmdstan()
library("fitdistrplus")
library("performance")
library("parallel")
library("parallelly")
library("broom.mixed")
library("tidybayes")
library("plotly")
library("viridis")
library("tidyverse")
```

@@

summary(pointsPerSeason_nullModel)
summary(pointsPerSeason_nullModel)$r.squared
AIC(pointsPerSeason_nullModel)
MuMIn::AICc(pointsPerSeason_nullModel)
```

@@

We use the `emmeans` package [@R-emmeans] to determine the model-implied number of points per season by position.

@@

summary(pointsPerSeason_position)
MuMIn::r.squaredGLMM(pointsPerSeason_position)
emmeans::emmeans(pointsPerSeason_position, "positionFactor")
performance::icc(pointsPerSeason_position)
AIC(pointsPerSeason_position)
```

@@

summary(pointsPerSeason_positionAgeFixedLinearSlopes)
MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeFixedLinearSlopes)
emmeans::emmeans(pointsPerSeason_positionAgeFixedLinearSlopes, "positionFactor")
emmeans::emmeans(pointsPerSeason_positionAgeFixedLinearSlopes, "ageCentered20")
performance::icc(pointsPerSeason_positionAgeFixedLinearSlopes)
AIC(pointsPerSeason_positionAgeFixedLinearSlopes)
```

@@

summary(pointsPerSeason_positionAgeRandomLinearSlopes)
MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearSlopes)
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearSlopes, "positionFactor")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearSlopes, "ageCentered20")
performance::icc(pointsPerSeason_positionAgeRandomLinearSlopes)
AIC(pointsPerSeason_positionAgeRandomLinearSlopes)
```

@@

summary(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "positionFactor")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "ageCentered20")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "ageCentered20Quadratic")
performance::icc(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
AIC(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
```

@@

summary(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction, "positionFactor")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction, "ageCentered20")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "ageCentered20Quadratic")
performance::icc(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
AIC(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
```

@@

summary(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "positionFactor")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "ageCentered20")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "ageCentered20Quadratic")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "years_of_experience")
performance::icc(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
AIC(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
```

@@

summary(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
MuMIn::r.squaredGLMM(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
emmeans::emmeans(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "positionFactor")
emmeans::emmeans(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "ageCentered20")
emmeans::emmeans(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "ageCentered20Quadratic")
emmeans::emmeans(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "years_of_experience")
performance::icc(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
AIC(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)

>>>>

```{r}
library("petersenlab")
library("lme4")
library("lmerTest")
library("MuMIn")
library("emmeans")
library("sjstats")
library("mgcv")
library("AICcmodavg")
library("bbmle")
library("rstan")
library("brms")
library("cmdstanr") # todo: install.packages("cmdstanr", repos = c("https://stan-dev.r-universe.dev", getOption("repos"))); cmdstanr::check_cmdstan_toolchain(); cmdstanr::install_cmdstan()
library("fitdistrplus")
library("performance")
library("parallel")
library("parallelly")
library("broom.mixed")
library("tidybayes")
library("plotly")
library("viridis")
library("tidyverse")
```

@@

summary(pointsPerSeason_nullModel)

summary(pointsPerSeason_nullModel)$r.squared
AIC(pointsPerSeason_nullModel)
MuMIn::AICc(pointsPerSeason_nullModel)
```

@@

To obtain standardized regression coefficients, we use the `lm.beta.lmer()` function of the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab].
We use the `emmeans` package [@R-emmeans] to determine the model-implied number of points per season by position.

@@

summary(pointsPerSeason_position)
petersenlab::lm.beta.lmer(pointsPerSeason_position)

MuMIn::r.squaredGLMM(pointsPerSeason_position)
emmeans::emmeans(pointsPerSeason_position, "positionFactor")
performance::icc(pointsPerSeason_position)
AIC(pointsPerSeason_position)
```

@@

summary(pointsPerSeason_positionAgeFixedLinearSlopes)
petersenlab::lm.beta.lmer(pointsPerSeason_positionAgeFixedLinearSlopes)

MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeFixedLinearSlopes)
emmeans::emmeans(pointsPerSeason_positionAgeFixedLinearSlopes, "positionFactor")
emmeans::emmeans(pointsPerSeason_positionAgeFixedLinearSlopes, "ageCentered20")
performance::icc(pointsPerSeason_positionAgeFixedLinearSlopes)
AIC(pointsPerSeason_positionAgeFixedLinearSlopes)
```

@@

summary(pointsPerSeason_positionAgeRandomLinearSlopes)
petersenlab::lm.beta.lmer(pointsPerSeason_positionAgeFixedLinearSlopes)

MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearSlopes)
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearSlopes, "positionFactor")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearSlopes, "ageCentered20")
performance::icc(pointsPerSeason_positionAgeRandomLinearSlopes)
AIC(pointsPerSeason_positionAgeRandomLinearSlopes)
```

@@

summary(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
petersenlab::lm.beta.lmer(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)

MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "positionFactor")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "ageCentered20")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "ageCentered20Quadratic")
performance::icc(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
AIC(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
```

@@

summary(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
petersenlab::lm.beta.lmer(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)

MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction, "positionFactor")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction, "ageCentered20")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "ageCentered20Quadratic")
performance::icc(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
AIC(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
```

@@

summary(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
petersenlab::lm.beta.lmer(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)

MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "positionFactor")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "ageCentered20")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "ageCentered20Quadratic")
emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "years_of_experience")
performance::icc(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
AIC(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
```

@@

summary(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
petersenlab::lm.beta.lmer(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)

MuMIn::r.squaredGLMM(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
emmeans::emmeans(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "positionFactor")
emmeans::emmeans(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "ageCentered20")
emmeans::emmeans(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "ageCentered20Quadratic")
emmeans::emmeans(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "years_of_experience")
performance::icc(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
AIC(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
